

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Machine Learning Algorithms

## INDEX
#### 1.Linear regression for Prediction of Suicide count from 2013-2033
#### 2.With respect to Causes,Year        
#### 3.State,Year  
#### 4.Age grp,Year             
#### 5.Wrt State,Age grp,Year Culminated Model


### 0)Libraries,Extraction , cleaning and separating:
```{r,echo=TRUE}

df<-read.table('C:/Personal/Aravinda Stuff/Sem 5th/DV and FDA proj/Suicides in India 2001-2012.csv',header=T, sep=",")
#head(df)

#install.packages('tidyverse')
#install.packages('caret')
#install.packages("plotly")
library(tidyverse)
library(caret)
library(dplyr)
library(ggplot2)
library(plotly)
library(glmnet)

#Data Cleaning
#Replacing Values for UT
df["State"][df["State"] == "A & N Islands"]<-"A & N Islands-Ut"
df["State"][df["State"] == "Chandigarh"]<-"Chandigarh-Ut"
df["State"][df["State"] == "D & N Haveli"]<-"D & N Haveli-Ut"
df["State"][df["State"] == "Daman & Diu"]<-"Daman & Diu-Ut"
df["State"][df["State"] == "Lakshadweep"]<-"Lakshadweep-Ut"
df["State"][df["State"] == "Delhi"] <-"Delhi-Ut"
#head(df)
#Renaming causes
df["Type"][df["Type"]=="Bankruptcy or Sudden change in Economic"]<-"Sudden change in Economic Status or Bankruptcy"
df["Type"][df["Type"]=="By Other means (please specify)"]<-"By Other means"
df["Type"][df["Type"]=="Not having Children(Barrenness/Impotency"]<-"Not having Children(Impotency)"
df["Type"][df["Type"]=="By Jumping from (Building)"]<-"By Jumping from Building"
df["Type"][df["Type"]=="Hr. Secondary/Intermediate/Pre-Universit"]<-"Hr. Secondary/Intermediate/Pre-University"
df["Type"][df["Type"]=="Failure in Examination"]<-"Examination Failure"
df["Type"][df["Type"]=="By coming under running vehicles/trains"]<-"By road or railway accidents" 
df["Type"][df["Type"]=="Bankruptcy or Sudden change in Economic Status"]<-"Sudden change in Economic Status or Bankruptcy"
df["Type"][df["Type"]=="Not having Children (Barrenness/Impotency"]<-"Not having Children(Impotency)"
#causescount
#head(df)

#drop the unwanted State-titles
df1 <- df[!(df$State=="Total (Uts)" | df$State=="Total (All India)" |  df$State=="Total (States)"),]
#drop the values ==0 under Total
df2 <- df1[!(df1$Total==0),]
# drop the unwanted Types
df2 <- df2[!(df2$Type=="By Other means" | df2$Type=="Others (Please Specify)" | df2$Type=="Causes Not known" |  df2$Type=="Other Causes (Please Specity)"),]

#Spliting the dataframe into smaller dataframe based on the column "Type_code"
causesdf=filter(df2,df2$Type_code=="Causes")
edudf=filter(df2,df2$Type_code=="Education_Status")
meansdf=filter(df2,df2$Type_code=="Means_adopted")
professionaldf=filter(df2,df2$Type_code=="Professional_Profile")
socialdf=filter(df2,df2$Type_code=="Social_Status")

head(df2)
```



### 1.1)Linear Regression for prediction for suicide count from 2013-2033
```{r, echo=TRUE}
# Extracting the needed year and suicide count columns
suicide_count_overyears <- df2 %>% group_by(Year) %>% summarise(total_case=sum(Total))
suicide_count_overyears

#Testing co relation
cor(suicide_count_overyears$Year,suicide_count_overyears$total_case)
cor.test(suicide_count_overyears$Year,suicide_count_overyears$total_case)

#Partitioning into train and test
set.seed(123)
train_samples <- suicide_count_overyears$Year %>%
  createDataPartition(p=0.65,list=FALSE)
#train_samples
head(train_samples)
train <- suicide_count_overyears[train_samples,]
test <- suicide_count_overyears[-train_samples,]
#train
#test
#suicide_count_overyears

# LR Model creation
model <- lm(total_case~Year,data=train)
summary(model)

#Make predictions
pred <- model %>%
  predict(test)

pred

#Verification with actual value and predicted values
RMSE <- RMSE(pred,test$total_case)
RMSE

R2 <- R2(pred,test$total_case)
R2

hist(model$residuals)
qqnorm(model$residuals,ylab = "Residuals")
qqline(model$residuals)

# Prediction
future_years=data.frame(Year=c(2013:2033))
future_years$total_case <- model %>%
  predict(future_years)

future_years


# Bar plot for rise in suicide cases
total_suicide_count <- rbind(suicide_count_overyears,future_years)
#total
fig <- plot_ly(
  x = total_suicide_count$Year,
  y = total_suicide_count$total_case,
  name = "Variation of suicide count over the years 2013-2032",
  type = "bar",

)
fig <- fig %>% layout(title = "Suicide Trend Over the years",
                      barmode = 'group',
                      xaxis = list(title = "Years"),
                      yaxis = list(title = "Count"))

fig
```

### 1.2)Logistic Regression for prediction for suicide count from 2013-2033
```{r, echo=TRUE}

# Extracting the needed year and suicide count columns
suicide_count_overyears <- df2 %>% group_by(Year) %>% summarise(total_case=sum(Total))
suicide_count_overyears

#Testing co relation
cor(suicide_count_overyears$Year,suicide_count_overyears$total_case)
cor.test(suicide_count_overyears$Year,suicide_count_overyears$total_case)

#Partitioning into train and test
set.seed(123)
train_samples <- suicide_count_overyears$Year %>%
  createDataPartition(p=0.65,list=FALSE)
#train_samples
head(train_samples)
train <- suicide_count_overyears[train_samples,]
test <- suicide_count_overyears[-train_samples,]
train
test

as.factor(suicide_count_overyears$Total)


# Training model
logistic_modelLR <- glm(total_case~Year, data = train)
logistic_modelLR
pred <- logistic_modelLR %>%
  predict(test)
pred

RMSE <- RMSE(pred,test$total_case)
RMSE
R2 <- R2(pred,test$total_case)
R2
test
```

## 2)CAUSES Filtration

```{r}

topcauses<-causesdf%>%select(Type,Year,Total) %>% group_by(Type)%>% 
  summarise(Total=sum(Total)) %>% arrange(desc(Total))%>% head(10)
topcauses<-as.data.frame(topcauses)
head(topcauses)

topcauses1<-causesdf%>%select(Type,Year,Total) %>% group_by(Year,Type="Family Problems")%>%
  summarise(Total=sum(Total)) %>% arrange(desc(Type))
head(topcauses1)

topcauses2<-causesdf%>%select(Type,Year,Total) %>% group_by(Year,Type="Other Prolonged Illness")%>%
  summarise(Total=sum(Total)) %>% arrange(desc(Type))
head(topcauses2)

topcauses3<-causesdf%>%select(Type,Year,Total) %>% group_by(Year,Type="Insanity/Mental Illness")%>%
  summarise(Total=sum(Total)) %>% arrange(desc(Type))
head(topcauses3)

topcauses4<-causesdf%>%select(Type,Year,Total) %>% group_by(Year,Type="Love Affairs")%>%
  summarise(Total=sum(Total)) %>% arrange(desc(Type))
head(topcauses4)

topcauses5<-causesdf%>%select(Type,Year,Total) %>% group_by(Year,Type="Sudden change in Economic Status or Bankruptcy")%>%summarise(Total=sum(Total)) %>% arrange(desc(Type))
head(topcauses5)


top5causes=rbind(topcauses1,topcauses2,topcauses3,topcauses4,topcauses5)


```

## 2.1)CAUSES MLR

```{r}
set.seed(123)
train_samples <- top5causes$Year %>%
  createDataPartition(p=0.70,list=FALSE)

train <- top5causes[train_samples,]
test <- top5causes[-train_samples,]


# LR Model creation
model <- lm(Total~Year+Type,data=train)
summary(model)

#Make predictions
pred <- model %>%
  predict(test)

pred

#Verification with actual value and predicted values
RMSE <- RMSE(pred,test$Total)
RMSE

R2 <- R2(pred,test$Total)
R2

hist(model$residuals)
qqnorm(model$residuals,ylab = "Residuals")
qqline(model$residuals)

# Prediction
new.speeds <- data.frame(
  Year = c(2023, 2024, 2025) , Type = c("Family Problems","Love Affairs","Other Prolonged Illness") 
)
#(agedf)
predict(model, newdata = new.speeds)



future_years




```

## 2.2)Causes Lasso
```{r}
#Testing co relation
#cor(suicide_count_overyears$Year,suicide_count_overyears$total_case)
#cor.test(suicide_count_overyears$Year,suicide_count_overyears$total_case)

#Partitioning into train and test
set.seed(123)
train_samples <- top5causes$Total %>%
  createDataPartition(p=0.80,list=FALSE)

train <- top5causes[train_samples,]
test <- top5causes[-train_samples,]
#agedf

#install.packages("glmnet")


#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(data.matrix(train[, c('Year','Type')]), train$Total, alpha = 0.5)
cv_model
#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda

#[1] best_lambda=546.3268

#produce plot of test MSE by lambda value
plot(cv_model)


#Best Lasso model
#t=data.matrix(train[, c('Year','State','Age_group')])
#t
best_model <- glmnet(data.matrix(train[, c('Year','Type')]), train$Total, alpha = 0.5, lambda = best_lambda)
coef(best_model)

#Prediction

#define new observation
#new = matrix(c(2015,'Maharashtra',"45-59"), nrow=1, ncol=3) 
#data.matrix(c(2015,'Maharashtra',"45-59"))
#new
#use lasso regression model to predict response value
#predict(best_model, s = best_lambda, newx = new)

x=data.matrix(test[, c('Year','Type')])
#x
y=test$Total
#Metrics
y_predicted <- predict(best_model, s = best_lambda, newx = x)


RMSE <- RMSE(y,y_predicted)
RMSE

R2 <- R2(y,y_predicted)
R2

#find SST and SSE
#sst <- sum((y - mean(y))^2)
#sse <- sum((y_predicted - y)^2)

#find R-Squared
#rsq <- 1-sse/sst
#rsq

```


## 2.3)SVM for Causes
```{r}
library(e1071)
set.seed(123)

train_samples <- top5causes$Total %>% createDataPartition(p=0.65,list=FALSE)
train <- top5causes[train_samples,]
train$Type<-as.factor(train$Type)
test <- top5causes[-train_samples,]
test$Type<-as.factor(test$Type)
#agedf

# MLR Model creation
causessvm <- svm(Total~Year+Type,data=train, kernel = 'linear')
summary(causessvm)

#Make predictions
pred <- causessvm %>%predict(test)
pred

#Verification with actual value and predicted values
RMSE <- RMSE(pred,test$Total)
RMSE

R2 <- R2(pred,test$Total)
R2



```

## 2.4)Logistic Regression for Causes
```{r}
#Partitioning into train and test
set.seed(123)
train_samples <- top5causes$Total %>%
  createDataPartition(p=0.80,list=FALSE)

train <- top5causes[train_samples,]
test <- top5causes[-train_samples,]


as.factor(top5causes$Total)

# Training model
logistic_modelC <- glm(Total~Year+Type, data = train)
logistic_modelC
pred <- logistic_modelC %>%
  predict(test)
pred

RMSE <- RMSE(pred,test$Total)
RMSE
R2 <- R2(pred,test$Total)
R2
test
```

## 3)Top 3 states filtration
```{r,echo=TRUE}
 
topstate<-df2%>%filter(!State %in% c("Total (All India)","Total (States)","Total (Uts)"))%>%select(State,Year,Total) %>% group_by(State)%>% 
  summarise(Total=sum(Total)) %>% arrange(desc(Total))%>% head(10)
topstate<-as.data.frame(topstate)
topstate

topstate1<-df2%>%filter(!State %in% c("Total (All India)","Total (States)","Total (Uts)"))%>%
  select(State,Year,Total) %>% group_by(Year,State="Maharashtra")%>%
  summarise(Total=sum(Total)) %>% arrange(desc(State))
topstate2<-df2%>%filter(!State %in% c("Total (All India)","Total (States)","Total (Uts)"))%>%
  select(State,Year,Total) %>% group_by(Year,State="West Bengal")%>%
  summarise(Total=sum(Total)) %>% arrange(desc(State))
topstate3<-df2%>%filter(!State %in% c("Total (All India)","Total (States)","Total (Uts)"))%>%
  select(State,Year,Total) %>% group_by(Year,State="Andhra Pradesh")%>%
  summarise(Total=sum(Total)) %>% arrange(desc(State))

top3state=rbind(topstate1,topstate2,topstate3)
#View(top3state)
```

## 3.1)TOP 3 STATES MLR
```{r}

# Extracting the needed year and suicide count columns

#Testing co relation
#cor(suicide_count_overyears$Year,suicide_count_overyears$total_case)
#cor.test(suicide_count_overyears$Year,suicide_count_overyears$total_case)

#Partitioning into train and test
set.seed(123)
train_samples <- top3state$Year %>% createDataPartition(p=0.80,list=FALSE)
train <- top3state[train_samples,]
test <- top3state[-train_samples,]


# LR Model creation
model <- lm(Total~Year+State,data=train)
summary(model)

#Make predictions
pred <- model %>%
  predict(test)

pred

#Verification with actual value and predicted values
RMSE <- RMSE(pred,test$Total)
RMSE

R2 <- R2(pred,test$Total)
R2

hist(model$residuals)
qqnorm(model$residuals,ylab = "Residuals")
qqline(model$residuals)

# Prediction
new.speeds <- data.frame(
  Year = c(2013, 2014, 2025) , State = c("Maharashtra","West Bengal","Maharashtra"))
#(agedf)
predict(model, newdata = new.speeds)

#future_years=data.frame(Year=c(2013:2033))
#future_years$total_case <- model %>%
#  predict(future_years)





```

## 3.2)Lasso for top 3 states

```{r, echo=TRUE}
#install.packages("glmnet")
library(glmnet)

#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(data.matrix(top3state[, c('Year','State')]), top3state$Total, alpha = 1)
cv_model
#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda

#[1] best_lambda=1123.63

#produce plot of test MSE by lambda value
plot(cv_model)


#Best Lasso model
best_model <- glmnet(data.matrix(top3state[, c('Year','State')]), top3state$Total, alpha = 1, lambda = best_lambda)
coef(best_model)

#Prediction

#define new observation
#new = matrix(c(2005,'Maharashtra'), nrow=1, ncol=2) 
#new
#use lasso regression model to predict response value
#predict(best_model, s = best_lambda, newx = new)

x=data.matrix(top3state[, c('Year','State')])
y=top3state$Total
#Metrics
y_predicted <- predict(best_model, s = best_lambda, newx = x)

#find SST and SSE
sst <- sum((y - mean(y))^2)
sse <- sum((y_predicted - y)^2)

#find R-Squared
rsq <- 1 - sse/sst
rsq
```


## 3.3)SVM for top 3 states
```{r}
top3state
library(e1071)
set.seed(123)

train_samples <- top3state$Total %>% createDataPartition(p=0.80,list=FALSE)
train <- top3state[train_samples,]
test <- top3state[-train_samples,]

#agedf

# MLR Model creation
statemodelsvm <- svm(Total~State,data=train, kernel = 'linear')
summary(statemodelsvm)

#Make predictions
pred <- statemodelsvm %>%predict(test)
pred

#Verification with actual value and predicted values
RMSE <- RMSE(pred,test$Total)
RMSE

R2 <- R2(pred,test$Total)
R2




```

## 3.4)Logistic Regression for top 3 states
```{r, echo=TRUE}
# Splitting dataset
set.seed(123)
train_samples <- top3state$Year %>% createDataPartition(p=0.65,list=FALSE)
#train_samples
head(train_samples)
train <- top3state[train_samples,]
test <- top3state[-train_samples,]
train
test

as.factor(top3state$Total)

# Training model
logistic_model <- glm(Total~Year+State, data = train)
logistic_model 
pred <- logistic_model %>%
  predict(test)
pred

RMSE <- RMSE(pred,test$Total)
RMSE
R2 <- R2(pred,test$Total)
R2
test
```

## 4)WRT AGE GRP Filtration
```{r}
agedf<-df2%>% select(Year,Age_group,Total)%>% 
  filter(!Age_group=="0-100")%>% 
  filter(!Age_group=="0-100+")%>%
  group_by(Year,Age_group)%>% 
  summarise(Total=sum(Total))%>% arrange(desc(Age_group))
agedf<-as.data.frame(agedf)
```

## 4.1)AGE GROUP MLR
```{r}

# Extracting the needed year and suicide count columns

#Testing co relation
#cor(suicide_count_overyears$Year,suicide_count_overyears$total_case)
#cor.test(suicide_count_overyears$Year,suicide_count_overyears$total_case)

#Partitioning into train and test
set.seed(123)
train_samples <- agedf$Year %>% createDataPartition(p=0.65,list=FALSE)
train <- agedf[train_samples,]
test <- agedf[-train_samples,]


# LR Model creation
model <- lm(Total~Year+Age_group,data=train)
summary(model)

#Make predictions
pred <- model %>%
  predict(test)

pred

#Verification with actual value and predicted values
RMSE <- RMSE(pred,test$Total)
RMSE

R2 <- R2(pred,test$Total)
R2

hist(model$residuals)
qqnorm(model$residuals,ylab = "Residuals")
qqline(model$residuals)

# Prediction
new.speeds <- data.frame(
  Year = c(2013, 2013, 2013) , Age_group = c("30-44","45-59","0-14"))
#(agedf)
predict(model, newdata = new.speeds)

#future_years=data.frame(Year=c(2013:2033))
#future_years$total_case <- model %>%
#  predict(future_years)





```

## 4.2)SVM Model for Age wrt Total

```{r}
library(e1071)
set.seed(123)

train_samples <- agedf$Total %>% createDataPartition(p=0.80,list=FALSE)
train <- agedf[train_samples,]
test <- agedf[-train_samples,]

#agedf

# MLR Model creation
modelsvm1 <- svm(Total~Year+Age_group,data=train, kernel = 'linear')
summary(modelsvm1)

#Make predictions
pred <- modelsvm1 %>%predict(test)
pred

#Verification with actual value and predicted values
RMSE <- RMSE(pred,test$Total)
RMSE

R2 <- R2(pred,test$Total)
R2

#find SST and SSE
sst <- sum((y - mean(y))^2)
sse <- sum((y_predicted - y)^2)

#find R-Squared
rsq <- 1-sse/sst
rsq


```


## 4.3)Lasso for age grp

```{r, echo=TRUE}
#install.packages("glmnet")
library(glmnet)

#Partitioning into train and test
set.seed(123)
train_samples <- agedf$Total %>%
  createDataPartition(p=0.80,list=FALSE)
train <- agedf[train_samples,]
test <- agedf[-train_samples,]

#agedf

#install.packages("glmnet")
#library(glmnet)

#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(data.matrix(train[, c('Year','Age_group')]), train$Total, alpha = 0.5)
cv_model
#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda

#[1] best_lambda=2565.932

#produce plot of test MSE by lambda value
plot(cv_model)


#Best Lasso model
#t=data.matrix(train[, c('Year','State','Age_group')])
#t
best_model <- glmnet(data.matrix(train[, c('Year','Age_group')]), train$Total, alpha = 0.5, lambda = best_lambda)
coef(best_model)

#Prediction

#define new observation
#new = matrix(c(2015,'Maharashtra',"45-59"), nrow=1, ncol=3) 
#data.matrix(c(2015,'Maharashtra',"45-59"))
#new
#use lasso regression model to predict response value
#predict(best_model, s = best_lambda, newx = new)

x=data.matrix(test[, c('Year','Age_group')])
#x
y=test$Total
#Metrics
y_predicted <- predict(best_model, s = best_lambda, newx = x)


RMSE <- RMSE(y,y_predicted)
RMSE

R2 <- R2(y,y_predicted)
R2

#find SST and SSE
#sst <- sum((y - mean(y))^2)
#sse <- sum((y_predicted - y)^2)

#find R-Squared
#rsq <- 1-sse/sst
#rsq
```
## 4.4)Logistic Regression for age grp

```{r, echo=TRUE}
#Partitioning into train and test
set.seed(123)
train_samples <- agedf$Year %>% createDataPartition(p=0.65,list=FALSE)
train <- agedf[train_samples,]
test <- agedf[-train_samples,]
train
test

as.factor(agedf$Total)

# Training model
logistic_modelAG <- glm(Total~Year+Age_group, data = train)
logistic_modelAG
pred <- logistic_modelAG %>%
  predict(test)
pred

RMSE <- RMSE(pred,test$Total)
RMSE
R2 <- R2(pred,test$Total)
R2
test
```

## 5)Wrt State,Age grp,Year Culminated Model Filtration
```{r}
####
# Extracting the needed year and suicide count columns
topstate1<-df2%>%filter(!State %in% c("Total (All India)","Total (States)","Total (Uts)"))%>%
  select(State,Year,Age_group,Total) %>%
  filter(!Age_group=="0-100")%>% 
  filter(!Age_group=="0-100+")%>%
  group_by(Year,Age_group,State="Maharashtra")%>%
  summarise(Total=sum(Total)) %>% arrange(desc(State))

topstate2<-df2%>%filter(!State %in% c("Total (All India)","Total (States)","Total (Uts)"))%>%
  select(State,Year,Age_group,Total) %>%
  filter(!Age_group=="0-100")%>% 
  filter(!Age_group=="0-100+")%>%
  group_by(Year,Age_group,State="West Bengal")%>%
  summarise(Total=sum(Total)) %>% arrange(desc(State))

topstate3<-df2%>%filter(!State %in% c("Total (All India)","Total (States)","Total (Uts)"))%>%
  select(State,Year,Age_group,Total) %>%
  filter(!Age_group=="0-100")%>% 
  filter(!Age_group=="0-100+")%>%
  group_by(Year,Age_group,State="Andhra Pradesh")%>%
  summarise(Total=sum(Total)) %>% arrange(desc(State))
#topstate3
model=rbind(topstate1,topstate2,topstate3)
modellasso=rbind(topstate1,topstate2,topstate3)
modelsvm=rbind(topstate1,topstate2,topstate3)
modelLog=rbind(topstate1,topstate2,topstate3)

```




## 5.1)Wrt State,Age grp,Year Culminated Model Lasso
```{r}
#Testing co relation
#cor(suicide_count_overyears$Year,suicide_count_overyears$total_case)
#cor.test(suicide_count_overyears$Year,suicide_count_overyears$total_case)

#Partitioning into train and test
set.seed(123)
train_samples <- modellasso$Total %>%
  createDataPartition(p=0.80,list=FALSE)
train <- modellasso[train_samples,]
test <- modellasso[-train_samples,]
#agedf

#install.packages("glmnet")
library(glmnet)

#perform k-fold cross-validation to find optimal lambda value
cv_model <- cv.glmnet(data.matrix(train[, c('Year','State','Age_group')]), train$Total, alpha = 0)
cv_model
#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min


best_lambda

#[1] best_lambda=2565.932

#produce plot of test MSE by lambda value
plot(cv_model)


#Best Lasso model
#t=data.matrix(train[, c('Year','State','Age_group')])
#t
best_model <- glmnet(data.matrix(train[, c('Year','State','Age_group')]), train$Total, alpha = 0, lambda = best_lambda)
coef(best_model)

#Prediction

#define new observation
#new = matrix(c(2015,'Maharashtra',"45-59"), nrow=1, ncol=3) 
#data.matrix(c(2015,'Maharashtra',"45-59"))
#new
#use lasso regression model to predict response value
#predict(best_model, s = best_lambda, newx = new)

x=data.matrix(test[, c('Year','State','Age_group')])
#x
y=test$Total
#Metrics
y_predicted <- predict(best_model, s = best_lambda, newx = x)

y_predicted

RMSE <- RMSE(y,y_predicted)
RMSE

R2 <- R2(y,y_predicted)
R2

#find SST and SSE
#sst <- sum((y - mean(y))^2)
#sse <- sum((y_predicted - y)^2)
#sst
#sse
#find R-Squared
#rsq <- 1-sse/sst
#rsq

```

## 5.2)SVM Wrt State,Age grp,Year Culminated Model

```{r}
library(e1071)
set.seed(123)
train_samples <- modelsvm$Total %>%
  createDataPartition(p=0.80,list=FALSE)

train <- modelsvm[train_samples,]
test <- modelsvm[-train_samples,]
#agedf

# MLR Model creation
modelsvm <- svm(Total~Year+Age_group+State,data=train, kernel = 'linear')
summary(modelsvm)

#Make predictions
pred <- modelsvm %>%predict(test)
pred

#Verification with actual value and predicted values
RMSE <- RMSE(pred,test$Total)
RMSE

R2 <- R2(pred,test$Total)
R2

```


## 5.3)LOGISTIC REGRESSION State,Age grp,Year Culminated Model
```{r}
##BIG MODEL LOGISTIC REGRESSION
# Splitting dataset
set.seed(123)
train_samplesbm <- modelLog$Year %>% createDataPartition(p=0.65,list=FALSE)
#train_samples
head(train_samplesbm)
train <- modelLog[train_samplesbm,]
test <- modelLog[-train_samplesbm,]
#train
#test

as.factor(modelLog$Total)

# Training model
logistic_modelbm <- glm(Total~Year+Age_group+State, data = train)
logistic_modelbm 
pred <- logistic_modelbm %>%
  predict(test)
pred
RMSE <- RMSE(pred,test$Total)
RMSE
R2 <- R2(pred,test$Total)
R2
#test
```

## 5.4)Wrt State,Age grp,Year Culminated Model MLR
```{r}
#Testing co relation
#cor(suicide_count_overyears$Year,suicide_count_overyears$total_case)
#cor.test(suicide_count_overyears$Year,suicide_count_overyears$total_case)

#Partitioning into train and test
set.seed(123)
train_samples <- model$Total %>%
  createDataPartition(p=0.80,list=FALSE)
train <- model[train_samples,]
test <- model[-train_samples,]

#agedf

# MLR Model creation
model <- lm(Total~Year+Age_group+State,data=train)
summary(model)

#Make predictions
pred <- model %>%
  predict(test)

pred

#Verification with actual value and predicted values
RMSE <- RMSE(pred,test$Total)
RMSE

R2 <- R2(pred,test$Total)
R2

hist(model$residuals)
qqnorm(model$residuals,ylab = "Residuals")
qqline(model$residuals)

# Prediction
new.speeds <- data.frame(
  Year = c(2013, 2013, 2013,2022,2023,2024) , Age_group = c("30-44","45-59","0-14","15-29","60+","30-44") ,State=c("Maharashtra","West Bengal","Andhra Pradesh","Andhra Pradesh","Maharashtra","West Bengal")
)
#(agedf)
predict(model, newdata = new.speeds)

```
